---
title: 'Reproducible Research: Peer Assessment 1'
author: "Sherby509"
date: "2/23/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## load libraries
```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
```
## process datasets
```{r}
fileUrl<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
     download.file(fileUrl, destfile="StormData.csv.bz2", method="curl")
```
## load the dataset
```{r}
stormData1 <- read.csv("StormData.csv.bz2")
```
## structure oh the data
```{r}
str(stormData1)
```
## Defining variables
```{r}
stormD <- select(stormData1, BGN_DATE, EVTYPE, PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP, FATALITIES, INJURIES)
```
## Format the BGN_DATE variable as a date

```{r}
stormD$BGN_DATE <- as.Date(stormD$BGN_DATE, "%m/%d/%Y")
```
```{r}
stormD$YEAR <- year(stormD$BGN_DATE)
```
## Considering events since 1996
```{r}
stormD <- filter(stormD, YEAR >= 1996)
```
## Considering events with health impact or economic damage
```{r}
stormD <- filter(stormD, PROPDMG > 0 | CROPDMG > 0 | FATALITIES > 0 | INJURIES > 0)
```
```{r}
table(stormD$PROPDMGEXP)
```
```{r}
table(stormD$CROPDMGEXP)
```
```{r}
stormD$PROPDMGEXP <- toupper(stormD$PROPDMGEXP)
```
```{r}
stormD$CROPDMGEXP <- toupper(stormD$CROPDMGEXP)
```
```{r}
stormD$CROPDMGFACTOR[(stormD$CROPDMGEXP == "")] <- 10^0
```
```{r}
stormD$CROPDMGFACTOR[(stormD$CROPDMGEXP == "?")] <- 10^0
```
```{r}
stormD$CROPDMGFACTOR[(stormD$CROPDMGEXP == "0")] <- 10^0
```
```{r}
stormD$CROPDMGFACTOR[(stormD$CROPDMGEXP == "2")] <- 10^2
```
```{r}
stormD$CROPDMGFACTOR[(stormD$CROPDMGEXP == "K")] <- 10^3
```
```{r}
stormD$CROPDMGFACTOR[(stormD$CROPDMGEXP == "M")] <- 10^6
```
```{r}
stormD$CROPDMGFACTOR[(stormD$CROPDMGEXP == "B")] <- 10^9
```

```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "")] <- 10^0
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "-")] <- 10^0
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "?")] <- 10^0
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "+")] <- 10^0
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "0")] <- 10^0
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "1")] <- 10^1
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "2")] <- 10^2
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "3")] <- 10^3
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "4")] <- 10^4
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "5")] <- 10^5
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "6")] <- 10^6
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "7")] <- 10^7
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "8")] <- 10^8
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "H")] <- 10^2
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "K")] <- 10^3
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "M")] <- 10^6
```
```{r}
stormD$PROPDMGFACTOR[(stormD$PROPDMGEXP == "B")] <- 10^9
```
```{r}
stormD <- mutate(stormD, HEALTHIMP = FATALITIES + INJURIES)
```
```{r}
stormD <- mutate(stormD, ECONOMICCOST = PROPDMG * PROPDMGFACTOR + CROPDMG * CROPDMGFACTOR)
```
```{r}
stormD$EVTYPE <- toupper(stormD$EVTYPE)
```
```{r}
dim(data.frame(table(stormD$EVTYPE)))
```
```{r}
healthImpact <- with(stormD, aggregate(HEALTHIMP ~ EVTYPE, FUN = sum))
```
```{r}
subset(healthImpact, HEALTHIMP > quantile(HEALTHIMP, prob = 0.95))
```
```{r}
stormD$EVTYPE[(stormD$EVTYPE == "TSTM WIND")] <- "THUNDERSTORM WIND"
```
```{r}
stormD$EVTYPE[(stormD$EVTYPE == "HURRICANE/TYPHOON")] <- "HURRICANE (TYPHOON)"
```
```{r}
economicCost <- with(stormD, aggregate(ECONOMICCOST ~ EVTYPE, FUN = sum))
```
```{r}
subset(economicCost, ECONOMICCOST > quantile(ECONOMICCOST, prob = 0.95))
```
```{r}
stormD$EVTYPE[(stormD$EVTYPE == "HURRICANE")] <- "HURRICANE (TYPHOON)"
```
```{r}
stormD$EVTYPE[(stormD$EVTYPE == "STORM SURGE")] <- "STORM SURGE/TIDE"
```

```{r}
healthImpact <- stormD %>% 
                group_by(EVTYPE) %>% 
                summarise(HEALTHIMP = sum(HEALTHIMP)) %>% 
                arrange(desc(HEALTHIMP))
```
```{r}
gr1 <- ggplot(healthImpact[1:10,], aes(x=reorder(EVTYPE, -HEALTHIMP),y=HEALTHIMP,color=EVTYPE)) + 
      geom_bar(stat="identity", fill="purple") + 
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
      xlab("Kind of Event") + ylab("Number of fatalities & injuries") +
      theme(legend.position="none") +
      ggtitle("Weather Events that cause Death and Injuries in the US")
```
```{r}
print(gr1)
```
## The chart shows that Tornados are the most fatalities weather events for people’s health.
```{r}
economicCost <- stormD %>% 
                group_by(EVTYPE) %>% 
                summarise(ECONOMICCOST = sum(ECONOMICCOST)) %>% 
                arrange(desc(ECONOMICCOST))
```
```{r}
gr1 <- ggplot(economicCost[1:10,], aes(x=reorder(EVTYPE, -ECONOMICCOST),y=ECONOMICCOST,color=EVTYPE)) + 
      geom_bar(stat="identity", fill="grey") + 
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
      xlab("Kind of Event") + ylab("Economic Damage cost  (Billion $$$)") +
      theme(legend.position="none") +
      ggtitle("Weather Events with the Greatest Economic Cost")
```
## print the chart
```{r}
print(gr1)
```
# The chart shows that Floods have the biggest economical cost.
