---
title: "Reproductible Research (Project 1)"
author: "Sherby509"
date: "2/25/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
getwd()
```

## Let's load the data (1)
```{r}
activity <- read.csv("activity.csv")
```
```{r}
activity <- read.csv("activity.csv")
```

```{r}
library(ggplot2)
```

```{r}
library(dplyr)
```

```{r}
Sys.setlocale("LC_TIME", "English")
```
```{r}
str(activity)
```
Now we see thee type of variable:
Steps
Date
Interval

## Let's calculate the number of steps by day
```{r}
Steps.Day <- aggregate(activity$steps, list(activity$date), FUN=sum)
```
```{r}
colnames(Steps.Day) <- c("Date", "Steps")
```
## lets print it
```{r}
print(Steps.Day)
```
## Let draw the histogram (2)
```{r}
g <- ggplot(Steps.Day, aes(Steps))
```
```{r}
g+geom_histogram(boundary=0, binwidth=2500, col="darkgreen", fill="purple")+ggtitle("Histogram of Number of Steps per Day")+xlab("Steps")+ylab("Frequency")+theme(plot.title = element_text(face="bold", size=12))+scale_x_continuous(breaks=seq(0,25000,2500))+scale_y_continuous(breaks=seq(0,18,2))
```
## Let's find the mean and the median (3)

# The mean
```{r}
mean(Steps.Day$Steps, na.rm=TRUE)
```
# The median
```{r}
median(Steps.Day$Steps, na.rm=TRUE)
```
## Now Time series plot of the average number of steps taken (4)
```{r}
Steps.Time <- aggregate(steps~interval,data=activity,FUN=mean,na.action=na.omit)
```
```{r}
Steps.Time$time <- Steps.Time$interval/100
```
## Lets draw the plot
```{r}
h <- ggplot(Steps.Time, aes(time, steps))
```
```{r}
h+geom_line(col="purple")+ggtitle("Average Number of Steps Taken interval")+xlab("Time")+ylab("Steps")+theme(plot.title = element_text(face="bold", size=20))
```
## The 5 minutes interval that contain the max of steps (5)
```{r}
Max.Steps <- tbl_df(Steps.Time)
```
```{r}
Max.Steps %>% select(time, steps) %>% filter(steps==max(Max.Steps$steps))
```
## Imputing missing Data
```{r}
MV <- tbl_df(activity)
```
```{r}
MV %>% filter(is.na(steps)) %>% summarize(missing_values = n())
```
## Replace the missing data
```{r}
activity$CompleteSteps <- ifelse(is.na(activity$steps), round(Steps.Time$steps[match(activity$interval, Steps.Time$interval)],0), activity$steps)

```
```{r}
activityFull <- data.frame(steps=activity$CompleteSteps, interval=activity$interval, date=activity$date)
```
```{r}
head(activityFull, n=10)
```
## Let's make the Histogram the max of steps taken each day with missing data filled in
```{r}
Steps.Day <- aggregate(activityFull$steps, list(activityFull$date), FUN=sum)
colnames(Steps.Day) <- c("Date", "Steps")
```
```{r}
g <- ggplot(Steps.Day, aes(Steps))
```
```{r}
g+geom_histogram(boundary=0, binwidth=2500, col="red", fill="blue")+ggtitle("Histogram of Steps Taken")+xlab("Steps")+ylab("Frequency")+theme(plot.title = element_text(face="bold", size=12))+scale_x_continuous(breaks=seq(0,25000,2500))+scale_y_continuous(breaks=seq(0,26,2))
```
## Let's find the new mean and the median (different from the fist one) 
# the mean
```{r}
mean(Steps.Day$Steps)
```
# the median
```{r}
median(Steps.Day$Steps)
```
## Are there differences in activity patterns between weekdays and weekends
# Create variable with date in correct format
```{r}
activityFull$RealDate <- as.Date(activityFull$date, format = "%Y-%m-%d")
```
```{r}
activityFull$weekday <- weekdays(activityFull$RealDate)
```
# create a new variable indicating weekday or weekend
```{r}
activityFull$DayType <- ifelse(activityFull$weekday=='Saturday' | activityFull$weekday=='Sunday', 'weekend','weekday')
```
# The first 15 values
```{r}
head(activityFull, n=15)
```
## Let's make a panel plot with a time series plot of the 5-minute interval as (x-axis) and  weekday days or weekend days as (y-axis). 
```{r}
Steps.TimeDT <- aggregate(steps~interval+DayType,data=activityFull,FUN=mean,na.action=na.omit)
```
```{r}
Steps.TimeDT$time <- Steps.Time$interval/100
```
```{r}
PL <- ggplot(Steps.TimeDT, aes(time, steps))
```
```{r}
PL+geom_line(col="purple")+ggtitle("Average Steps of 5 Minutes Interval: weekdays vs. weekends")+xlab("Time")+ylab("Steps")+theme(plot.title = element_text(face="bold", size=12))+facet_grid(DayType ~ .)
```

